### 2020-02-17 ordinal/ordinal.Rmd
ordinal tutorial

```{r init, include=F,echo=F}
knitr::opts_chunk$set(dev = 'svg', comment = NA, fig.path='Figs/',fig.width=8, fig.height=6,
                      echo=TRUE, warning = FALSE, error = FALSE, message = FALSE,
                      cache.path = '~/windows/temp/ordinal', cache=T)

sdt_ml_plot = function(model, adata, sdata){
    sdt = !is.null(adata$stimulus)
    o = optimizing(stan_model(model_code = model), data = sdata, init = 0)
    mp = matrix(o$par[grep('multinomial_p', names(o$par))], nrow = nrow(adata$data))
    obs = t(apply(adata$counts, 1, function(x)(x / sum(x))))
    if(sdt)
        stim = as.factor(adata$stimulus)
    f = as.factor(adata$data[,1])
    if(ncol(adata$data) > 1)
        for(i in 2:ncol(adata$data))
            f = f:as.factor(adata$data[,i])
    if(sdt){
        ggplot(data.frame(f = rep(f, ncol(mp)), stim = rep(stim, ncol(mp)), th = rep(1:ncol(mp), each = length(f)),
                          mp = as.vector(mp), obs = as.vector(obs)),
               aes(th, obs, group = stim, color = stim)) +
        geom_line(aes(y = mp)) +
        geom_point() +
        labs(title = 'Subj x Condition panels, lines = model, points = observed response distribution') +
        xlab('response = decision + rating') +
        ylab('p(response)') +
        facet_wrap(~ f)
    }else{
        ggplot(data.frame(f = rep(f, ncol(mp)), th = rep(1:ncol(mp), each = length(f)),
                          mp = as.vector(mp), obs = as.vector(obs)),
               aes(th, obs)) +
        geom_line(aes(y = mp)) +
        geom_point() +
        labs(title = 'lines = fitted model, points = observed response distribution') +
        xlab('response = decision + rating') +
        ylab('p(response)') +
        facet_wrap(~ f)
    }
}
```

# Hierarchical ordinal regression models in bhsdtr

This document contains a short tutorial on fitting hierarchical
ordinal polytomous regression models using the bhsdtr package, but
there is more; It also contains some explanations, illustrated with
annotated code, of several issues related to ordinal models in
general, and Signal Detection Theory and Item Response Theory in
particular. If you are interested in SDT, IRT, or ordinal models in
general, perhaps you will find something useful, new, or even mildly
provocative here.

Anyway, it is now possible to fit hierarchical (more than two possible
values) ordinal regression models in bhsdtr by using order-preserving
link functions for the thresholds. If you have a bunch of Likert-type
items, or confidence ratings, or PAS ratings, or anything of this sort
in your data, then this model may suit your needs. To my knowledge
this is a new kind of hierarchical ordinal model that can be used to
estimate variability in the thresholds in a way that until recently
was not possible.

Ordinal polytomous variables are common in psychology, they are
usually analyzed by fitting a linear model, and all sorts of bad
things happen because of this, but only a few people seem to care. I
for one do not always care about this. We will not address this kind
of problems here, because we will analyze ordinal data using
(hierarchical) ordinal models.

We will use the following libraries and global variables:

```{r echo=T,results="show"}
knitr::opts_chunk$set(echo=TRUE,results="show")
library(ggplot2)
library(bhsdtr)
library(Hmisc)
library(coda)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
## This variable indicates if the fitted models can be loaded from disk
fresh_start = FALSE
sumfun = function(x)c(quantile(x, c(.025, .975)), mean(x))[c(1,3,2)]
```

First, we will consider an ordinal model which assumes that there is
an underlying normally distributed latent variable (e.g., an evidence
distribution in Signal Detection Theory, or a latent trait in an IRT
model) and when the value of this variable is above the latent
threshold *i*, but below the threshold *i + i*, the response *i + 1*
is generated. Threshold *0* is fixed at *-infinity* and threshold *K*
is fixed at *+infinity*, which means that there are *K-1* "regular"
thresholds and *K* possible responses. Since once we know the total
number of observations and the number of occurrences of *K - 1*
different responses, the number of occurrences of the only remaining
kind of response is known, there are *K - 1* degrees of freedom in
such data. The response variable is ordinal, and so the thresholds are
ordered. This is how this model looks like when the thresholds are
equidistant and centered at the mean of the latent variable
distribution (they do not have to be equidistant nor do they have to
be centered):

```{r echo=F}
x = seq(-3, 3, length.out = 1000)
y = dnorm(x)
p = ggplot(data.frame(x = x, y = y), aes(x, y)) + geom_line(lty = 2)
crit = c(-2, -1, 0, 1, 2)
for(i in 1:length(crit))
    p = p + eval(substitute(geom_vline(aes(xintercept = cr)), list(cr = crit[i])))
p = p + scale_x_continuous(breaks = crit, labels = 1:length(crit), name = 'threshold / criterion')
for(i in 1:(length(crit) + 1))
    p = p + eval(substitute(geom_text(aes(x, y, label = label, size = 20)),
                            list(x = c(-3, crit)[i] + .5, y = max(y) / 2, label = paste(i))))
p
```

The number of parameters in this model matches the degrees of freedom
in the data, which means that the model is saturated, which means that
it always fits the data, but it does not mean that there is no
uncertainty associated with the estimates of its parameters.

We will be using this dataset...

```{r echo=T,results="show"}
data(gabor)
```

... from which we will select two subsets of data for one condition,
one subset (*d1*) with only the "noise" trials and the other (*d2*)
with both the "noise" and the "signal" stimuli. We will also calculate
the response variable by combining the information about the stimulus,
the accuracy, and the rating.

```{r echo=T,results="show"}
## r = 1 ~ absolutely clear image left, r = 4 ~ no experience left, r
## = 5 ~ no experience right, ... r = 8 ~ absolutely clear image right
gabor$r = combined_response(gabor$stim, gabor$rating, gabor$acc)
d1 = gabor[(gabor$duration == '32 ms') & (gabor$order == 'DECISION-RATING') & (gabor$stim == 0),]
d2 = gabor[(gabor$duration == '32 ms') & (gabor$order == 'DECISION-RATING'),]
```

We cannot naively model the effects in the thresholds, because the
thresholds have to stay ordered, and assuming, for example, that each
threshold has a normal distribution due to the effect of participants
would violate the ordering assumption.

To supplement our model with a hierarchical linear regression
structure the vector of the thresholds has to be translated to the
unconstrained vector gamma. The 'log_distance' order-preserving link
function is one solution to this problem, but other order-preserving
link functions are also implemented in the bhsdtr package. The choice
of the link function may sometimes be important, but, since we will
not introduce numerical predictors for the thresholds, we can choose
any link function which seems most convenient.

The log_distance function is convenient for our purposes because it is
simple, and because the elements of the gamma vector under this link
function are easy to interpret and may be directly related to simple
research questions. In particular, one of the elements of the gamma
vector under the log_distance link function, i.e., the middle
threshold, will soon become important. This link function represents
the main threshold/criterion directly as a free parameter (here
gamma[4], because K = 8, and 8 / 2 = 4); Every other criterion is
represented as log of distance between the criterion and the one
criterion adjacent to it which is closer to the main criterion. This
way the order is preserved, and each threshold *i* is associated with
the corresponding unconstrained element *i* of the gamma
vector. Random and fixed effects in the thresholds are represented as
(hierarchical) linear effects in gamma. The two representations are
isomorphic, which means that we can always translate between them if
we need to.

```{r echo=T,results="show"}        
link = 'log_distance'
```

To fit a model we have to aggregate the data (this makes the process
of sampling much faster), create the fixed- and random-effects lists,
and create the data structures and the model code required by
stan. Note that here we introduce a new parameter - eta - which
represents the average internal value. If we are going to model all
the thresholds (gamma parameter vector) then this parameter is
redundant, because a constant shift in the thresholds is equivalent to
the opposite effect in eta. The bhsdtr package will automatically fix
the first element (here the only element) of the eta fixed effects
vector at 0.

First, we will fit a model which assumes that the pattern of the
thresholds is the same for all the participants, but the mean of the
latent distribution may differ across participants. If this makes
little sense to you please take a look at the picture of the ordinal
model with ordered thresholds and imagine that the underlying
distribution has a different mean for different participants.

Note that the two assumptions, i.e., that the thresholds are constant
across participants and latent means or samples are participant-, not
item-specific seem to be accepted in the great majority of
applications of Item Response Theory. This is reflected in the IRT
terminology: thresholds are often called "item-parameters" and latent
means or samples are often called "person-parameters". Some even say
that IRT can provide separate estimates of the true item- and
person-parameters.

```{r echo=T,results="show"}
fixed = list(eta = ~ 1, gamma = ~ 1)
random = list(list(group = ~ id, eta = ~ 1))
adata = aggregate_responses(d1, response = 'r', variables = 'id')
model = make_stan_model(random, model = 'ordinal', gamma_link = link)
sdata = make_stan_data(adata, fixed, random, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit = stan(model_code = model, data = sdata,
                   pars = c('eta_fixed', 'gamma_fixed', 'eta_random_1', 'eta_sd_1'),
                   init_r = .5,
                   iter = 5000,
                   chains = 4)
    saveRDS(fit, '~/windows/temp/ordinal_fit.rds')
}else{
    fit = readRDS('~/windows/temp/ordinal_fit.rds')
}
print(fit, probs = c(.025, .975), pars = c('eta_fixed', 'gamma_fixed'))
```

The neff and Rhat columns look good. Here is how the model fits (this
is a maximum likelihood fit):

```{r echo=T,results="show"}
sdt_ml_plot(model, adata, sdata)
```

So how do the "item-parameters" look like according to this model?

```{r echo=T,results="show"}
crit = gamma_to_crit(as.data.frame(fit))
round(t(apply(crit, 2, sumfun)), 2)
```

Note that the main criterion (gamma[4,1]) is *positive*, which makes
sense, since when the stimulus is of class "noise" the middle
criterion / threshold / average position of the thresholds is *to the
right* of the evidence distribution mean. Otherwise, as we will see
later, in this particular case, and perhaps also in many other similar
cases, these thresholds-as-"item-parameters" are probably meaningless.

Now we will introduce the model which accounts for the
inter-individual differences in the pattern of the thresholds by
modelling the gamma random effects. We cannot simultaneously model the
eta random effects in this model, since all the information that can
be captured by the eta parameter (i.e., the constant shift in the
thresholds) is already captured by the fixed and the random gamma effects.

```{r echo=T,results="show"}                
adata0 = aggregate_responses(d1, response = 'r', variables = c('id'))
## The only element of the eta fixed effects vector will be automatically fixed at 0
fixed0 = list(eta = ~ 1, gamma = ~ 1)
## Each participant has a unique patterh of the thresholds
random0 = list(list(group = ~ id, gamma = ~ 1))
model0 = make_stan_model(random0, model = 'ordinal', gamma_link = link)
sdata0 = make_stan_data(adata0, fixed0, random0, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit0 = stan(model_code = model0, data = sdata0,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit0, '~/windows/temp/ordinal_fit_0.rds')
}else{
    fit0 = readRDS('~/windows/temp/ordinal_fit_0.rds')
}
```

```{r echo=T,results="show"}              
print(fit0, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
## neff and Rhat look good
```

The neff and Rhat columns look good. Below we translate between the
gamma and the criteria vectors (fixed effects = sample average):

```{r echo=T,results="show"}                
s0 = as.data.frame(fit0)
crit0 = gamma_to_crit(s0, gamma_link = link)
```

Are the sample average thesholds different in the two models? Off
course they are, since the models are non-linear and averaging over
anything that affects any parameter of such models is a bad idea:

```{r echo=T,results="show"}                
crit_sm = as.data.frame(t(apply(crit, 2, sumfun)))
crit0_sm = as.data.frame(t(apply(crit0, 2, sumfun)))
crit_sm$model = 'fixed thresholds'
crit0_sm$model = "random thresholds"
crit_sm$threshold = 1:7 - .1
crit0_sm$threshold = 1:7 + .1
df = rbind(crit_sm, crit0_sm)
names(df)[1:3] = c('lo', 'mean', 'hi')
ggplot(df, aes(threshold, mean, group = model, color = model)) + geom_point() + geom_line() + geom_errorbar(aes(ymin = lo, ymax = hi))
```

The point and interval estimates of the thresholds based on the
thresholds-as-"item-parameters" model are severely biased. In
particular, the uncertainty in the estimates is severely
underestimated in this model. Was it obvious to you that the model
with no threshold random effects is such a bad idea?

Now we will fit a model which is similar to an SDT model, but there
will be two important differences. We will not assume that the answer
"signal" cannot be less likely for the stimulus "signal" than for the
stimulus "noise", i.e., we will not assume that d' must be
non-negative, and we will *not* assume that the pattern of the
thresholds is the same for "noise" and "signal" stimuli. If you know
your Signal Detection Theory well, this may seem to you like a
perverse idea. If that is the case, please bear with me.

In this model each participant may have two different sets of
thresholds, one for "noise" and one for "signal" stimuli. Since the
thresholds are represented as random effects, we will account for the
fact that "random" thresholds may be correlated, and they are - if
some threshold is higher (lower) than the sample average, then the
threshold above (below) it may also be higher (lower).

Here we introduce the effect of the stimulus on the gamma / thresholds
vector. We use separate intercepts and slopes parametrization (-1 +
stim), i.e., the thresholds are to be estimated for each stimulus class
instead of estimating them for "noise" stimuli and estimating the
difference in the thresholds between "signal" and "noise".

```{r echo=T,results="show"}                
fixed1 = list(eta = ~ 1, gamma = ~ -1 + stim)
random1 = list(list(group = ~ id, gamma = ~ -1 + stim))
## We have to make sure that stim is a factor for the separate
## intercepts and slopes parametrization to work as intended.
d2$stim = as.factor(d2$stim)
## Note that because for an ordinal model the stimulus variable (here
## the 'stim' variable) is not special, we have to include it in the
## additional variables vector
adata1 = aggregate_responses(d2, response = 'r', variables = c('id', 'stim'))
sdata1 = make_stan_data(adata1, fixed1, random1, model = 'ordinal', gamma_link = link)
model1 = make_stan_model(random1, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit1 = stan(model_code = model1, data = sdata1,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit1, '~/windows/temp/ordinal_fit_1.rds')
}else{
    fit1 = readRDS('~/windows/temp/ordinal_fit_1.rds')
}
## We will not look at the eta parameter since it was fixed at 0
print(fit1, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
## neff and Rhat look good
```

Now we can translate between the gamma and the criteria vectors (fixed
effects = sample average) for each stimulus:

```{r echo=T,results="show"}
s1 = as.data.frame(fit1)
crit1 = gamma_to_crit(s1, 1, gamma_link = link)
crit2 = gamma_to_crit(s1, 1, beta_index = 2, gamma_link = link)
sumfun = function(x)c(quantile(x, c(.025, .975)), mean(x))[c(1,3,2)]
crit1_sm = apply(crit1, 2, sumfun)
crit2_sm = apply(crit2, 2, sumfun)
```

The criteria for the stimulus class "noise" look like this:

```{r echo=T,results="show"}
round(t(crit1_sm), 2)
```

and the criteria for the stimulus class "signal" look like this:

```{r echo=T,results="show"}
round(t(crit2_sm), 2)
```

Note the difference between the main criterion (row 4). This
difference approximates d'. Even though, as we will later see, this
model does not correctly account for the effect of the stimulus, we can
still recover this effect pretty well, because d' is equivalent to a
constant shift of all the thresholds in one predefined
direction. Shifting all the criteria to the right is equivalent to
shifting the underlying distribution of "latent values" (or evidence)
to the left, so we have to subtract the average position of the
criteria for the "signal" stimuli from the average position of the
criteria for the "noise" stimuli. Here is our first attempt:

```{r echo=T,results="show"}
dprim1 = sumfun(s1[, 'gamma_fixed[4,1]'] - s1[, 'gamma_fixed[4,2]'])
round(dprim1, 2)
```

We will later see that this is our sample average *d'*, but it also
isn't.

In an SDT model we assume that the stimulus class affects *only* the
d' parameter, and that both criteria (gamma) and *d'* may differ
between the participants (or items, etc.). In bhsdtr d' is modelled as
log(d') = delta, which forces it to be non-negative. This is an
important assumption of Signal Detection Theory - if this assumption
is violated, for example, because of the response reversal in some
participants or conditions, then an SDT model is not valid and a
non-trivial generalization of SDT should be used instead (see the
bhsdtr preprint for a more detailed explanation and some
references). Otherwise, this is the same general hierarchical ordinal
regression model.

```{r echo=T,results="show"}   
fixed2 = list(delta = ~ 1, gamma = ~ 1)
random2 = list(list(group = ~ id, delta = ~ 1, gamma = ~ 1))
```

Because of the importance of the stimulus class ("noise", or 1, and
"signal", or 2) in SDT models we have to indicate which variable
represents this information ( aggregate_responses(..., stimulus =
'stim', ...) ).

```{r echo=T,results="show"}
adata2 = aggregate_responses(d2, stimulus = 'stim', response = 'r', variables = c('id'))
sdata2 = make_stan_data(adata2, fixed2, random2, model = 'sdt', gamma_link = link)
model2 = make_stan_model(random2, model = 'sdt', gamma_link = link)
if(fresh_start){
    fit2 = stan(model_code = model2, data = sdata2,
                pars = c('delta_fixed', 'gamma_fixed', 'delta_sd_1', 'gamma_sd_1',
                         'delta_random_1', 'gamma_random_1',
                         ## We will need this for the plot_sdt_fit function
                         'counts_new'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit2, '~/windows/temp/ordinal_fit_2.rds')
}else{
    fit2 = readRDS('~/windows/temp/ordinal_fit_2.rds')
}
print(fit2, probs = c(.025, .975), pars = c('delta_fixed', 'gamma_fixed', 'delta_sd_1', 'gamma_sd_1'))
## neff and Rhat look good
```

So how well do the two d' estimates match? We can translate between
the delta and the d' parameters by using exponentiation. We have to do
this first, before doing anything else with the posterior samples,
because the exponential function is non-linear.

```{r echo=T,results="show"}
s2 = as.data.frame(fit2)
dprim2 = sumfun(exp(s2[,'delta_fixed[1,1]']))
round(cbind(dprim1, dprim2), 2)
```

The estimates match poorly. Or do they? What about the individual *d'*
estimates? To calculate the individual shifts in the thresholds /
criteria we have to add the random effects to the fixed effect:

```{r echo=T,results="show"}   
dprim1i = (s1[['gamma_fixed[4,1]']] + s1[, grep('gamma_random_1\\[.+,4,1]', names(s1))]) -
    (s1[['gamma_fixed[4,2]']] + s1[, grep('gamma_random_1\\[.+,4,2]', names(s1))])
dprim1i = as.data.frame(t(apply(dprim1i, 2, sumfun)))
dprim2i = as.data.frame(t(apply(exp(s2[['delta_fixed[1,1]']] +
                                    s2[, grep('delta_random_1', names(s2))]), 2, sumfun)))
names(dprim1i) = names(dprim2i) = c('lo', 'mean', 'hi')
dprim1i$model = 'ordinal'
dprim1i$i = 1:nrow(dprim1i) - .1
dprim2i$model = 'sdt'
dprim2i$i = 1:nrow(dprim2i) + .1
ggplot(rbind(dprim1i, dprim2i), aes(i, mean, group = model, color = model)) +
    geom_point() + geom_errorbar(aes(ymin = lo, ymax = hi)) +
    geom_abline(intercept = 0, slope = 0) +
    ylab('d\'') +
    labs(title = sprintf('Correlation between individual d\' estimates: r = %.2f',
                         cor(dprim1i$mean, dprim2i$mean)))
```

The point and interval estimates match quite well, with some
notable exceptions. According to the ordinal model some credible
intervals around individual d' point estimates include 0 and one point
estimate is even below 0, which contradicts an important assumption of
the SDT model. We will address this issue later.

If the participant-specific d' estimates match, the sample average d'
should also match, right? Maybe the SDT model does not fit the data
well?

```{r echo=T,results="show"}   
plot_sdt_fit(fit2, adata2, c('id'), type = 'response', bw = F, verbose = F)
```

The SDT model seems to fit very well in a sense that the observed
distributions of responses are well within the 95% posterior
predictive intervals for each participant. However, this does not mean
much; an SDT model with multiple criteria is not saturated, so it does
not have to fit the data, but it does not leave too many degrees of
freedom in the data, so we can safely assume a priori that it will
usually fit the data well, regardless if it is true or not.

The discrepancy between the sample average *d'* estimates from both
models is not really a consequence of allowing for negative *d'*
values in the ordinal model. This discrepancy is a consequence of the
assumed distribution of individual shifts / *d'* values. In the
ordinal model we assumed that the shifts (*= d'*) are normally
distributed, whereas in the SDT model we assumed that *log(d')* are
normally distributed. We can see if the assumption that *log(d')* (or
*d'*) are normally distributed is approximately correct in our
dataset. First, let's take a look at participant-specific delta
estimates in the SDT model:

```{r echo=T,results="show"}
ggplot(data.frame(delta = log(dprim2i$mean)), aes(sample = delta)) +
    stat_qq() +
    ylab('Sample quantile of log(d\') = delta random effects') +
    xlab('Theoretical normal quantile') +
    labs(title = 'SDT model')
```

The individual *delta = log(d')* estimates in the SDT model seem to be
approximately normally distributed. We cannot plot log(d') estimates
based on the ordinal model, because according to this model one
participant has a negative *d'* value. However, we can see if the *d'*
point estimates based on the ordinal model are approximately normally
distributed:

```{r echo=T,results="show"}   
ggplot(data.frame(dprim = dprim1i$mean), aes(sample = dprim)) +
    stat_qq() +
    ylab('Sample quantile of d\' random effects') +
    xlab('Theoretical normal quantile') +
    labs('Ordinal model')
```

They seem to be approximately normally distributed as well. The
average / fixed effect *d'* in each model is estimated as the mean of
different distributions: in the ordinal model participant-specific
shifts are assumed to be normally distributed and in the SDT model
delta = log(d') random effects are assumed to be normally
distributed. We can easily account for this by making use of the
relationship between the parameters of the normal distribution and the
associated log-normal distribution:

```{r echo=T,results="show"}   
round(cbind(sumfun(exp(s2[['delta_sd_1[1,1]']]^2 / 2 + s2[['delta_fixed[1,1]']])),
            dprim1), 2)
```

Now we see that the two models give very similar point and interval
estimates of sample-average *d'*.

To summarize, the SDT model is much simpler (it has only one set of
thresholds for each participant), it fits the data well at the
individual level, and it correctly assumes that the individual *d'* /
shift values are log-normally distributed. SDT is a clear winner,
right? Not yet it isn't.

An SDT model is not only a statistical model, it also has *causal*
assumptions. Because the stimulus class was chosen at random on every
trial, every estimate of the statistical effect of the stimulus class
on any parameter of the model is an unbiased estimate of the total
causal effect of the stimulus class on this parameter. Moreover, in
such designs every statistical effect of the participant factor is an
estimate of the property or the causal effect of the
participants. This means that both the individual *d'* parameters and
the individual thresholds are participant-, not item-, parameters in
this model. So far so good.

However, in the SDT model we have also assumed that the stimulus class
can *only* affect the distance between the evidence distribution means
(= *d'*), whereas in the ordinal model we allowed for the possibility
of the stimulus class having an effect on the thresholds. It may seem
odd to you to allow for this, but there is nothing in the design of
the study that justifies the causal assumption of selective
influence. The estimates of the thresholds are based on the data that
were generated after the stimulus was presented, and so,
theoretically, the *pattern* of the thresholds could be affected by
the stimulus class as well.

The selectivity assumption is central to Signal Detection Theory. If
you know your Signal Detection Theory well, please take a moment and
try to recall one instant of this important assumption being
explicitly addressed in a paper or a book on Signal Detection Theory.

It does not matter if an SDT model with selectivity assumption fits
the data well, because if the effect of the stimulus class is not
selective, the model is simply false. In particular, when the
selectivity assumption is false, an SDT model cannot deconfound
sensitivity / *d'* from bias / threshold / criteria effects.

Before we go any further, just for fun, let's see what happens if we
introduce yet another assumption of selective influence, one that is
often accepted in Item Response Theory. We will fit a hierarchical SDT
model in which we will call the thresholds "item-parameters" and we
will call *d'* a "person-parameter".

```{r echo=T,results="show"}
fixed3 = list(delta = ~ 1, gamma = ~ 1)
random3 = list(list(group = ~ id, delta = ~ 1))
adata3 = aggregate_responses(d2, stimulus = 'stim', response = 'r', variables = c('id'))
sdata3 = make_stan_data(adata3, fixed3, random3, model = 'sdt', gamma_link = link)
model3 = make_stan_model(random3, model = 'sdt', gamma_link = link)
if(fresh_start){
    fit3 = stan(model_code = model3, data = sdata3,
                pars = c('delta_fixed', 'delta_sd_1', 'gamma_fixed', 'delta_random_1',
                         ## We will need this for the plot_sdt_fit function
                         'counts_new'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit3, '~/windows/temp/ordinal_fit_3.rds')
}else{
    fit3 = readRDS('~/windows/temp/ordinal_fit_3.rds')
}
print(fit3, probs = c(.025, .975), pars = c('delta_fixed', 'gamma_fixed', 'delta_sd_1'))
## neff and Rhat look good
```

Does the model fit?

```{r echo=T,results="show"}
plot_sdt_fit(fit3, adata3, type = 'response', bw = F, verbose = F)
```

At the aggregated data level the fit is not perfect, but it is not
that terrible either. What about the fit at the individual level?

```{r echo=T,results="show"}
plot_sdt_fit(fit3, adata3, 'id', type = 'response', bw = F, verbose = F)
```

A few points are outside the predictive intervals, but maybe that's
ok? What if we compared the point and interval estimates of the so-called "person-parameters"?

```{r echo=T,results="show"}   
s3 = as.data.frame(fit3)
dprim3i = as.data.frame(t(apply(exp(s3[['delta_fixed[1,1]']] + s3[, grep('delta_random_1', names(s3))]), 2, sumfun)))
names(dprim3i) = c('lo', 'mean', 'hi')
dprim3i$model = 'person- and item-parameters'
dprim3i$i = 1:nrow(dprim3i) - .1
ggplot(rbind(dprim2i, dprim3i), aes(i, mean, group = model, color = model)) +
    geom_point() + geom_errorbar(aes(ymin = lo, ymax = hi)) +
    geom_abline(intercept = 0, slope = 0) +
    labs(title = sprintf('Correlation between individual d\' estimates: r = %.2f', cor(dprim2i$mean, dprim3i$mean)))    
```

This does not look good at all. In particular, some of the point and
interval estimates are clearly biased. If we compare the average width
of the intervals...

```{r echo=T, results="show"}
round(mean(dprim3i$hi - dprim3i$lo) / mean(dprim2i$hi - dprim2i$lo), 2)
```

... we see that the simplified model not only gives biased
"person-parameter" estimates but the estimates seem more reliable than
they truly are (here by about 20%). It does not make sense to
look at the estimates of the "item-parameters" since we already know
that they have little in common with the true thresholds.

Now back to Signal Detection Theory and the assumption of selective
influence. We can at least try to test this assumption. We will not
alter the ordinal model code to account for the non-negativity of d'
(and the log-normal distribution of random d' effects) just yet. As we
already know, even without this correction the estimates of
participants' shifts are similar to the *d'* estimates.

Note that we do not use the separate intercepts and slopes
parametrization here (i.e., ~ -1 + stim), because we are very much
interested in the *effect* of the stimulus class on the *pattern* of
the thresholds.

```{r echo=T,results="show"}   
fixed4 = list(eta = ~ 1, gamma = ~ stim)
random4 = list(list(group = ~ id, gamma = ~ stim))
adata4 = aggregate_responses(d2, response = 'r', variables = c('id', 'stim'))
sdata4 = make_stan_data(adata4, fixed4, random4, model = 'ordinal', gamma_link = link)
model4 = make_stan_model(random4, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit4 = stan(model_code = model4, data = sdata4,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit4, '~/windows/temp/ordinal_fit_4.rds')
}else{
    fit4 = readRDS('~/windows/temp/ordinal_fit_4.rds')
}
print(fit4, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
```

Here we see the estimated average shift / *d'* thought of as the mean
of normally distributed *d'* (not *delta = log(d')*) random effects
again:

```{r echo=T,results="show"}
s4 = as.data.frame(fit4)
round(cbind(sumfun(-s4[['gamma_fixed[4,2]']]), dprim1), 2)
```

Ok, we did simplify things significantly, but it still makes sense to
take a look at the other gamma effects, because these other effects
correspond to the variability in the *pattern* of the thresholds, and
the estimates of the thresholds in our ordinal model seem perfectly
fine.

```{r echo=T,results="show"}   
round(HPDinterval(as.mcmc(s4[, grep('gamma_fixed', names(s4))])), 2)
```

Among the parameters that represent the stimulus-induced average
differences between the gamma parameters (i.e., gamma_fixed[.,2]) the
gamma_fixed[4,2] parameter does not seem to be the only one that is
non-zero. This indicates that the pattern of the thresholds - not only
their average position - is affected by the stimulus
class. Interestingly, the random effects' standard deviations tell a
different story:

```{r echo=T,results="show"}   
round(HPDinterval(as.mcmc(s4[, grep('gamma_sd_1', names(s4))])), 2)
```

Note the clear pattern for the gamma_sd_1[.,2] parameters. Even though
the *sample-average* pattern of the thresholds seems to be different
for the two stimulus classes (gamma_fixed[.,2] except for
gamma_fixed[4,2] ~ d'), this difference in pattern seems to be fairly
constant across the participants: all the gamma_sd_1[.,2] parameters
seem to be near-zero (the 95% credible intervals contain zero), except
for the gamma_sd_1[4,2] parameter, which corresponds to the
variability in *d'* values. This is *really* weird, since the pattern
of the thresholds for the "noise" stimuli clearly varies with
participants - none of the 95% HPD intervals for the gamma_sd_1[.,1]
parameters include zero: in terms of standard deviations of posterior
distributions of standard deviations (phew) most are in fact quite far
from zero.

If only we could do a better job and use a model which respects the
non-negativity assumption. Is it really that complicated? Let's index
the rows of our dataset by *i*. Because of the way the model was
parametrized...

```{r echo=T,results="show"}   
head(sdata4$X_gamma)
```

when stim = "noise" = 0, gamma[4]_i = gamma_fixed[4,1] +
gamma_random_1[g_i][4,1], where g_i is the participant index in row
*i*. We can leave this part of the model intact, since this is the
main criterion position, which is an unconstrained parameter, but
also, as we can see below, the observed distribution of the main
criterion random effects is approximately normal:

```{r echo=T,results="show"}
ggplot(data.frame(gamma_4 = apply(s4[['gamma_fixed[4,1]']] + s4[, grep('gamma_random_1\\[.+,4,1]', names(s4))], 2, mean)),
       aes(sample = gamma_4)) +
    stat_qq() +
    ylab('Sample quantile of main criterion random effects') +
    xlab('Theoretical normal quantile') +
    labs('Ordinal model')
```

However, when stim = "signal" = 1, gamma[4]_i = gamma_fixed[4,1] +
gamma_random_1[g][4,1] + gamma_fixed[4,2] + gamma_random_1[g][4,2]. We
want this other part (the last sum which represents the difference in
gamma[4] between the "noise" and "signal" stimuli) to represent log of
average distance between the thresholds = shift = *d'*, so we have to
subtract it from gamma[4]_i, exponentiate this sum, and add it again
to gamma[4]_i.

```{r echo=T,results="show"}   
fixed5 = list(eta = ~ 1, gamma = ~ stim)
random5 = list(list(group = ~ id, gamma = ~ stim))
adata5 = aggregate_responses(d2, response = 'r', variables = c('id', 'stim'))
sdata5 = make_stan_data(adata5, fixed5, random5, model = 'ordinal', gamma_link = link)
model5_ = make_stan_model(random5, model = 'ordinal', gamma_link = link)
## Here we introduce the necessary change in the model code. Kb2
## stands for K by 2 = the middle criterion index.
model5 = gsub('criteria\\[Kb2] = gamma\\[Kb2];',
              'if(X_gamma[n,2] == 0){
                  criteria[Kb2] = gamma[Kb2];
               }else{ 
                  criteria[Kb2] = gamma[Kb2] - (gamma_fixed[Kb2,2] + gamma_random_1[group_1[n]][Kb2,2]) - exp(gamma_fixed[Kb2,2] + gamma_random_1[group_1[n]][Kb2,2]);
               }',
              model5_)
```

This way the gamma[4,1] parameter represents the position of the main
criterion and gamma[4,2] = delta = log(d').

```{r echo=T,results="show"}   
if(fresh_start){
    fit5 = stan(model_code = model5, data = sdata5,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit5, '~/windows/temp/ordinal_fit_5.rds')
}else{
    fit5 = readRDS('~/windows/temp/ordinal_fit_5.rds')
}
print(fit5, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
```

Now let's have a look at the *d'* estimate (fixed effect) and compare
it with the *d'* estimate in the SDT model:

```{r echo=T,results="show"}
s5 = as.data.frame(fit5)
round(cbind(sumfun(exp(s5[['gamma_fixed[4,2]']])), dprim2), 2)
```

This is close. Not the same, but the model is different. What about
the bias = main criterion relative the the midpoint between the main
criteria?
```{r echo=T,results="show"}   
round(cbind(sumfun(s5[['gamma_fixed[4,1]']] - exp(s5[['gamma_fixed[4,2]']]) / 2),
      sumfun(s2[['gamma_fixed[4,1]']])), 2)
```

Also close.

We have fitted a hierarchical SDT model which respects the
non-negativity assumption, the ordering of the thresholds, accounts
for the log-normal distribution of the *d'* random effects, allows for
random effects in both the *d'* and the individual threshold
parameters while still respecting the SDT assumptions (thanks to an
order-preserving gamma link function and the delta log link function
which respects the non-negativity assumption), and accounts for the
correlations between the random effects. First such model was
described in the bhsdtr preprint. This model, however, does not assume
that the stimulus class affects only the *d'* parameter. We can now
again see if there is any evidence of non-selectivity:

```{r echo=T,results="show"}   
round(HPDinterval(as.mcmc(s5[, grep('gamma_fixed', names(s5))])), 2)
```

```{r echo=T,results="show"}                
round(HPDinterval(as.mcmc(s5[, grep('gamma_sd_1', names(s5))])), 2)
```

We see again that gamma_sd_1[.,2] parameters, except for
gamma_sd_1[4,2] = log(d'), may be zero - the effect of the stimulus on
all the log-distances seems to be more or less the same for each
participant. What is going on here?

Interaction is tricky, you see. For example, if you have an additive
effect of two factors on the original scale of the dependent
variable...

```{r echo=T,results="show",fig.asp=.6}   
df = expand.grid(f1 = 1:2, f2 = 1:2)
df$y = df$f1 + df$f2
df$f1 = as.factor(df$f1)
df$f2 = as.factor(df$f2)
ggplot(df, aes(f1, y, group = f2, color = f2)) + geom_line()
```

...then after applying a non-linear transformation to the dependent
variable you may obtain a significant interacive effect:

```{r echo=T,results="show",fig.asp=.6}
ggplot(df, aes(f1, exp(y), group = f2, color = f2)) + geom_line()
```

This means that whenever you see an interactive effect it may be an
artefact of the scale of the dependent variable! Every time I say it
it makes me feel uneasy. So, instead of examining the effects in the
gamma parameters. let's have a look at the average *criteria /
thresholds*. Notice that we have to add the effect on gamma to gamma
for stimulus = "noise" to obtain gamma for stimulus = "signal".

```{r echo=T,results="show"}   
crit1 = gamma_to_crit(s5, gamma_link = link)
crit2 = gamma_to_crit(s5[, grep('gamma_fixed\\[.+,1]', names(s5))] + s5[, grep('gamma_fixed\\[.+,2]', names(s5))], gamma_link = link)
```

Now we can directly test if the *pattern* of the criteria may be
different by comparing *centered* criteria vectors:
```{r echo=T,results="show"}
round(HPDinterval(as.mcmc((crit2 - apply(crit2, 1, mean)) - (crit1 - apply(crit1, 1, mean)))), 2)
```

When we center the criteria for the two classes of stimuli we no
longer see any evidence of the differences in their relative
positions. The only thing that seems to differ, and not by much, is
the main criterion, but this is confounded with *d'* in our model (I
think). Selectivity seems to hold, kind of.

This is how you *begin* to test a hierarchical SDT model. An SDT model
has other problematic assumptions. One is that the time-accuracy
tradeoff can be ignored. If participants or conditions differ in the
response time, then the differences in *d'* may not mean that
sensitivity is different, because participants may simply take more
time to respond in certain experimental conditions, which inflates the
*d'* estimate. You need something like the diffusion model to deal
with this problem. If you combine the diffusion model with ratings,
you also need something like a hierarchical diffusion model with
ratings and an order-preserving link function. Sounds like something
that could perhaps be implemented in the bhsdtr package in the future.

The other problems that need to be addressed have to do with
non-linearity and aggregation. If you fit an SDT model, or any other
non-linear model, to data aggregated over grouping factors, such as
participants or items, because of non-linearity, which does not play
well with aggregation, some or all of the point and interval estimates
will be biased, perhaps severely so, and the inference will be
invalid. You can see how horribly biased the point *and* interval
estimates of every SDT parameter may become when you overly aggregate
your data in the bhsdtr preprint. I am now in the process of examining
the direction and magnitude of bias in SDT models fitted to
overly-aggregated data by using the excellent Confidence database
created by Doby Rahnev.

All this means that every SDT analysis to this day that was done using
data aggregated over participants is invalid. In particular, to my
knowledge, to this day there is little *valid* evidence that the
Unequal Variance SDT model fits the data better than the Equal
Variance SDT model.

Note also that here we were aggregating over trials, which means that,
if the SDT parameters may differ between trials, and they certainly do
differ, the only question is by how much, then all the SDT estimates
that I have presented to you are biased to an unknown extent. There is
a lot to be done before we can say how well does the humble SDT model
deconfound sensitivity from bias in any particular case.

What if we tried this trick with testing the selectivity assumption,
but using the Unequal Variance SDT model, or the meta-d' model, or
some other generalization of the SDT model? We cannot introduce any
new parameters to our model, because it is already saturated. What we
can do, for example, is fit an Unequal Variance SDT model, estimate
the ratio of the variances, fix the ratio, introduce it to the
saturated ordinal model, thus turning it into the Fixed Unequal
Variance Ordinal Model. If there is evidence of non-selectivity in
this model, we can keep trying with other modifications. If
non-selectivity disappears, we have evidence that the selectivity
assumption may be correct *and* that the Unequal Variance SDT model
may be better than the Equal Variance SDT model.

# A rant about Item Response Theory

An SDT model is essentially an ordinal model with some additional
assumptions (e.g., selectivity, non-negativity of d', no time-accuracy
tradeoff issues, etc.). When there is more than one criterion /
threshold this model represents the distribution of polytomous ordinal
responses. Because by definition a polytomous item has a multinomial
distribution and the multinomial distribution does not have a separate
parameter for the residual variance, when item-parameter variability
is not accounted for by the model it can only be hidden or misplaced
(overdispersion). The correct way to account for item-parameter
variability in ordinal polytomous items is to allow for individual
threshold effects while respecting the ordering of the
thresholds. This can only be achieved by using the order-preserving
link functions.

Assume that the ordered responses ranging from 1 to K come from the
same general kind of process as the responses in an SDT model when the
stimulus is 1 ("noise"). For example, when John is faced with the
Likert-type item "How often did you cry during the last two weeks" an
unknown process produces some "internal experience" value. This latent
value manifests itself in the response "never" (= 1) if it is low
enough, "a few times" (= 2) if it is a bit higher than some latent
threshold, or in the response "frequently" (= 3) if it is higher than
some other, still higher, threshold. The thresholds are assumed to be
ordered, because otherwise, the item would not have the intended
meaning and thus would not be a valid instrument of measuring
depression.

Now consider a participant who repeatedly responds to the same item
and who has the memory of this event wiped out after each
response. Despite the wiping out of memory we can still expect some
variability in the responses due to the instability of the complex
process which produces the internal value, as well as due to the
instability of the decision criteria. This variability is certainly
there, the question is if it is non-negligible, and if so, what can be
done about it.

Unfortunately, without additional assumptions, it is impossible to
model both kinds of variability at the same time even in the
hypothetical repeated-measurement-with-memory-wipeout-scenario,
because every shift in the internal value is indistinguishable from a
uniform effect in the thresholds, i.e., a constant shift of all the
thresholds. As I hope you already realize, the required additional
assumptions that make such models identifiable under certain
conditions are necessarily *causal*.

In many situations, this is both a statistical and a conceptual
problem. In particular, when the design is purely observational -
e.g., questionnaire measurement - without additional causal
assumptions, the way of describing the process of responding to a
Likert-type item in terms of internal values and thresholds is simply
redundant. As far as the design of the study is concerned, all we can
say about the internal value is that it is relative to the thresholds
- and the position of each threshold is relative to the positions of
the remaining thresholds and to the position of the internal
value. Without the selectivity assumption, to say that John thinks
that he was not crying at all during the last two weeks and to say
that *John's* threshold for the response "a few times" is above the
amount of crying that John thinks he experienced, or decided to
reveal, is to say the same thing.

It follows that *in such situations, we do not need the internal value
parameter at all*, unless there is some factor about which we can
perhaps assume that it can *only* produce a uniform effect in the
thresholds, such as the stimulus class ("noise" or "signal") in a
binary classification task, which can *perhaps* be assumed to induce a
change (*d'*) in the internal value only; In such cases the internal
value parameter may be justified and useful because it does constrain
the model. We have obtained evidence that this assumption perhaps
holds for the analyzed dataset.

Notice that an SDT model is just an Item Response Theory model with
the cumulative distribution and a non-negative effect of the stimulus
class on the latent mean. So what about the central promise of Item
Response Theory?  Isn't it true that we can separate the internal
values ("person parameters") from the thresholds ("item parameters")
and place them on a common scale? Once the participants leave the room
the items retain their estimated "difficulty levels", right? The item
difficulty (=thresholds) is not defined relative to some sample of
participants who filled the questionnaire. How is this possible if the
two kinds of effects are completely confounded? In particular, how
does Item Response Theory achieve this magical separation when *each
person provides only a few responses, usually just one, to each item*?

It is usually not an *achievement* of Item Response Theory, it is just
an unrealistic *assumption*. The illusory nature of this achievement
becomes obvious as soon as we realize that without additional strong
causal assumptions it is impossible to distinguish between genuine
internal value effects and (dis)simulation (i.e., faking), which
conceptually corresponds to the change in the *participant*-specific,
not item-specific, decision thresholds, also known as
"item-parameters". There is nothing in the *design* of the process of
measurement by a questionnaire that makes it possible to distinguish
genuine differences in internal values from the differences in the way
of responding.

Because of random assignment of the stimulus class above chance
performance in a binary classification task cannot be faked, and so
the *d'* parameter definitely captures at least part of the effect of
the stimulus on the internal value. On the other hand, every possible
pattern of responses to a questionnaire can be faked. The only way to
distinguish the "item" and "person" effects is to *assume selective
influence*, i.e., that there is some factor f (such as the stimulus
class in a binary classification task, or, in case of a depression
questionnaire, participants trait, or a therapeutic intervention)
which can *only* affect the internal value, thus producing a
*constant* shift in the thresholds. The assumption of selective
influence makes the model with the two kinds of parameters
identifiable if there are repeated measurements for at least one
participant and one item or many participants provide single responses
to all of the items.

In the context of Item Response Theory the issue of measurement
non-invariance due to the variability in the "item-parameters" is
often reduced to the problem of Differential Item Functioning
(DIF). However, the variability in "item-parameters" that I talk about
here has nothing to do with DIF, because DIF is a *population* level
effect; An estimate of DIF is obtained when the estimates of
"item-parameters" obtained for two different (often large) samples of
participants are compared. However, these estimates are obtained by
*fixing the "item-parameters" within each sample*.

If we campared the estimates of the main decision criterion (i.e.,
bias) in SDT models fitted to two different large samples of
participants the difference would likely be negligible with very
narrow confidence / credible intervals. By the same token, instability
of item parameters in IRT models cannot be estimated by assuming
selective influence and comparing estimates of "item-parameters"
obtained in two different large samples, because large-sample average
estimates of item parameters are the best way to *hide or misplace*
(it becomes part of the variability in the "person-parameters") the
intra- or inter-individual variability in item parameters; By assuming
that internal values are participant-specific and item parameters are
item-specific we force the estimates of item parameters to become
estimates of average person-specific thresholds. This not only begs
the question, but also, because the model is non-linear, makes the
estimates of item-parameters *asymptotically biased*.

It does not take much thought to see that the assumption of constant
"item-parameters" at the group / sample / population level is utterly
unrealistic. It follows that the achievement of separating participant
and item effects in IRT models is often not an achievement at all, it
is just wishful thinking.

The justification of the selective influence assumption comes from a
psychological theory of the task, not from the design of the
study. All that is guaranteed by randomly assigning the stimulus class
(i.e., by choosing the stimulus class at random on every trial) is
that any statistical effect of the stimulus class is an unbiased
estimate of the total causal effect of the stimulus class. To
empirically justify the assumption of *selective* influence of the
stimulus class on the internal evidence value we have to *test* if the
stimulus class does not also affect the thresholds.

There are a few important advantages that an SDT model fitted to
binary classification task data has over an IRT model fitted to
questionnaire data. One is that the causal assumption of selective
influence of the stimulus class on the internal values is much more
plausible than the - implicit in most applications of IRT - causal
assumption of selective influence of participants on the
"person-parameters". The other is that in a typical binary
classification task case the process of responding is repeated many
times, the task is relatively simple ("was the gabor patch tilted left
or right" vs "do you wish you were as happy as other people seem to
be?") and the processes responsible for the generation of internal
values and for the placement of the decision thresholds are likely to
be more stable.

The thresholds are determined by the items, by the properties of the
participants, and by the context in which the measurement takes
place. The assumption that important latent aspects of the process of
psychological measurement are constant across participants is so
ridiculous it should be laughed at on the street.