### 2020-02-17 ordinal/Readme.Rmd
the third layer

---
title: "Ordinal bhsdtr tutorial"
output:
  md_document:
    variant: markdown_github
---
```{r init, include=F,echo=F}
knitr::opts_chunk$set(dev = 'svg', comment = NA, fig.path='Figs/',fig.width=8, fig.height=6,
                      echo=TRUE, warning = FALSE, error = FALSE, message = FALSE,
                      cache.path = '~/windows/temp/ordinal/', cache=T)

sdt_ml_plot = function(model, adata, sdata){
    sdt = !is.null(adata$stimulus)
    o = optimizing(stan_model(model_code = model), data = sdata, init = 0)
    mp = matrix(o$par[grep('multinomial_p', names(o$par))], nrow = nrow(adata$data))
    obs = t(apply(adata$counts, 1, function(x)(x / sum(x))))
    if(sdt)
        stim = as.factor(adata$stimulus)
    f = as.factor(adata$data[,1])
    if(ncol(adata$data) > 1)
        for(i in 2:ncol(adata$data))
            f = f:as.factor(adata$data[,i])
    if(sdt){
        ggplot(data.frame(f = rep(f, ncol(mp)), stim = rep(stim, ncol(mp)), th = rep(1:ncol(mp), each = length(f)),
                          mp = as.vector(mp), obs = as.vector(obs)),
               aes(th, obs, group = stim, color = stim)) +
        geom_line(aes(y = mp)) +
        geom_point() +
        labs(title = 'Subj x Condition panels, lines = model, points = observed response distribution') +
        xlab('response = decision + rating') +
        ylab('p(response)') +
        facet_wrap(~ f)
    }else{
        ggplot(data.frame(f = rep(f, ncol(mp)), th = rep(1:ncol(mp), each = length(f)),
                          mp = as.vector(mp), obs = as.vector(obs)),
               aes(th, obs)) +
        geom_line(aes(y = mp)) +
        geom_point() +
        labs(title = 'lines = fitted model, points = observed response distribution') +
        xlab('response = decision + rating') +
        ylab('p(response)') +
        facet_wrap(~ f)
    }
}
```

# Hierarchical ordinal regression models in bhsdtr

This document contains a short tutorial on fitting hierarchical
ordinal polytomous regression models using the bhsdtr package, but
there is more; It also contains some explanations, illustrated with
annotated code, of several issues related to ordinal models in
general, and Signal Detection Theory and Item Response Theory in
particular. If you are interested in any of these topics, perhaps you
will find something useful here.

Anyway, it is now possible to fit hierarchical (more than two possible
values) ordinal regression models in bhsdtr by using order-preserving
link functions for the thresholds. If you have a bunch of Likert-type
items, or confidence ratings, or PAS ratings, or anything of this sort
in your data, then this model may suit your needs. To my knowledge
this is a new kind of hierarchical ordinal model that can be used to
estimate variability in the thresholds in a way that until recently
was not possible.

Ordinal polytomous variables are common in psychology, they are
usually analyzed by fitting a linear model, and all sorts of bad
things happen because of this, but most people do not seem to care. I
for one do not always care about this. We will not address this kind
of problems here, because we will analyze ordinal data using
(hierarchical) ordinal models.

We will use the following libraries and global variables:

```{r echo=T,results="show"}
knitr::opts_chunk$set(echo=TRUE,results="show")
library(ggplot2)
library(bhsdtr)
library(Hmisc)
library(coda)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
## This variable indicates if the fitted models can be loaded from
## disk. I use it for caching
fresh_start = FALSE
sumfun = function(x)c(quantile(x, c(.025, .975)), mean(x))[c(1,3,2)]
```

First, we will consider an ordinal model which assumes that there is
an underlying normally distributed latent variable (e.g., an evidence
distribution in Signal Detection Theory, or a latent trait in an IRT
model) and when the value of this variable is above the latent
threshold *i*, but below the threshold *i + i*, the response *i + 1*
is generated. Threshold *0* is fixed at *-infinity* and threshold *K*
is fixed at *+infinity*, which means that there are *K-1* "regular"
thresholds and *K* possible responses. Since once we know the total
number of observations and the number of occurrences of *K - 1*
different responses, the number of occurrences of the only remaining
kind of response is known, there are *K - 1* degrees of freedom in
such data. The response variable is ordinal, and so the thresholds are
ordered. This is how this model looks like when the thresholds are
equidistant and centered at the mean of the latent variable
distribution (they do not have to be equidistant nor do they have to
be centered):

```{r echo=F}
x = seq(-3, 3, length.out = 1000)
y = dnorm(x)
p = ggplot(data.frame(x = x, y = y), aes(x, y)) + geom_line(lty = 2)
crit = c(-2, -1, 0, 1, 2)
for(i in 1:length(crit))
    p = p + eval(substitute(geom_vline(aes(xintercept = cr)), list(cr = crit[i])))
p = p + scale_x_continuous(breaks = crit, labels = 1:length(crit), name = 'threshold / criterion')
for(i in 1:(length(crit) + 1))
    p = p + eval(substitute(geom_text(aes(x, y, label = label, size = 20)),
                            list(x = c(-3, crit)[i] + .5, y = max(y) / 2, label = paste(i))))
p
```

The number of parameters in this model matches the degrees of freedom
in the data (if the data are aggregated), which means that the model
is saturated, which means that it always fits the data, but it does
not mean that there is no uncertainty associated with the estimates of
its parameters.

We will be using this dataset...

```{r echo=T,results="show"}
data(gabor)
```

... from which we will select two subsets of data for one condition,
one subset (*d1*) with only the "noise" trials and the other (*d2*)
with both the "noise" and the "signal" stimuli. We will also calculate
the response variable by combining the information about the stimulus,
the accuracy, and the rating.

```{r echo=T,results="show"}
## r = 1 ~ absolutely clear image left, r = 4 ~ no experience left, r
## = 5 ~ no experience right, ... r = 8 ~ absolutely clear image right
gabor$r = combined_response(gabor$stim, gabor$rating, gabor$acc)
d1 = gabor[(gabor$duration == '32 ms') & (gabor$order == 'DECISION-RATING') & (gabor$stim == 0),]
d2 = gabor[(gabor$duration == '32 ms') & (gabor$order == 'DECISION-RATING'),]
```

We cannot naively model the effects in the thresholds, because the
thresholds have to stay ordered, and assuming, for example, that each
threshold has a normal distribution due to the effect of participants
would violate the ordering assumption. To supplement our model with a
hierarchical linear regression structure the vector of the thresholds
has to be translated to the unconstrained vector gamma. The
'log_distance' order-preserving link function is one solution to this
problem, but other order-preserving link functions are also
implemented in the bhsdtr package. The choice of the link function may
sometimes be important, but, since we will not introduce quantitative
predictors for the thresholds, we can choose any link function which
seems most convenient.

The log_distance function is convenient for our purposes because it is
simple, and because the elements of the gamma vector under this link
function are easy to interpret and may be directly related to simple
research questions. This link function represents the main threshold /
criterion directly as a free parameter (here gamma[4], because K = 8,
and 8 / 2 = 4); Every other criterion is represented as log of
distance between the criterion and the one criterion adjacent to it
which is closer to the main criterion. This way the order is
preserved, and each threshold *i* is associated with the corresponding
unconstrained element *i* of the gamma vector. Random and fixed
effects in the thresholds are represented as (hierarchical) linear
effects in gamma. The two representations are isomorphic, which means
that we can always translate between the gamma and the ordered vectors
if we need to.

```{r echo=T,results="show"}        
link = 'log_distance'
```

To fit a model we have to aggregate the data as much as possible
without distorting the hierarchical structure (this makes the process
of sampling much faster), create the fixed- and random-effects lists,
and create the data structures and the model code required by
stan. Note that we introduce a new parameter - eta - which represents
the average internal value. If we are going to model the effects in
all the thresholds (gamma parameter vector) then this parameter is
redundant, because a constant shift in the thresholds is equivalent to
the opposite effect in eta. The bhsdtr package will automatically fix
the first element (here the only element) of the eta fixed effects
vector at 0, which will make this model identifiable.

First, we will fit the model which assumes that the pattern of the
thresholds is the same for all the participants, but the mean of the
latent distribution may differ across participants. If this makes
little sense to you please take a look at the picture of the ordinal
model with ordered thresholds and imagine that the underlying
distribution has a different mean for different participants.

Note that the two assumptions, i.e., that the thresholds are constant
across participants and that the latent means or samples are
participant-, not item-specific seem to be often accepted in
applications of Item Response Theory. This is also reflected in the
IRT terminology: thresholds are often called "item-parameters" and
latent means or samples are often called
"person-parameters".

```{r echo=T,results="show"}
fixed = list(eta = ~ 1, gamma = ~ 1)
random = list(list(group = ~ id, eta = ~ 1))
adata = aggregate_responses(d1, response = 'r', variables = 'id')
model = make_stan_model(random, model = 'ordinal', gamma_link = link)
sdata = make_stan_data(adata, fixed, random, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit = stan(model_code = model, data = sdata,
                   pars = c('eta_fixed', 'gamma_fixed', 'eta_random_1', 'eta_sd_1'),
                   init_r = .5,
                   iter = 5000,
                   chains = 4)
    saveRDS(fit, '~/windows/temp/ordinal_fit.rds')
}else{
    fit = readRDS('~/windows/temp/ordinal_fit.rds')
}
print(fit, probs = c(.025, .975), pars = c('eta_fixed', 'gamma_fixed'))
```

Here is how the model fits (this is a maximum likelihood fit):

```{r echo=T,results="show"}
## This function is not defined in the bhsdtr package
sdt_ml_plot(model, adata, sdata)
```

So how do the "item-parameters" look like according to this model?

```{r echo=T,results="show"}
crit = gamma_to_crit(as.data.frame(fit))
round(t(apply(crit, 2, sumfun)), 2)
```

Note that the main criterion (gamma[4,1]) is *positive*, which makes
sense, since when the stimulus is of class "noise" the middle
criterion / threshold / average position of the thresholds is *to the
right* of the evidence distribution mean. Otherwise, as we will see
later, in this particular case, and perhaps also in many other similar
cases, these thresholds-as-"item-parameters" are probably meaningless.

Now we will introduce the model which accounts for the
inter-individual differences in the pattern of the thresholds by
modelling the gamma random effects.

```{r echo=T,results="show"}                
adata0 = aggregate_responses(d1, response = 'r', variables = c('id'))
## The only element of the eta fixed effects vector will be automatically fixed at 0
fixed0 = list(eta = ~ 1, gamma = ~ 1)
## Each participant has a unique patterh of the thresholds / gamma
random0 = list(list(group = ~ id, gamma = ~ 1))
model0 = make_stan_model(random0, model = 'ordinal', gamma_link = link)
sdata0 = make_stan_data(adata0, fixed0, random0, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit0 = stan(model_code = model0, data = sdata0,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit0, '~/windows/temp/ordinal_fit_0.rds')
}else{
    fit0 = readRDS('~/windows/temp/ordinal_fit_0.rds')
}
```

```{r echo=T,results="show"}              
print(fit0, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
```

Below we translate between the gamma and the criteria vectors (fixed
effects):

```{r echo=T,results="show"}                
s0 = as.data.frame(fit0)
crit0 = gamma_to_crit(s0, gamma_link = link)
```

Are the average thesholds different in the two models? Off course they
are, since the models are non-linear and averaging over anything that
affects any parameter of such models is a bad idea:

```{r echo=T,results="show"}                
crit_sm = as.data.frame(t(apply(crit, 2, sumfun)))
crit0_sm = as.data.frame(t(apply(crit0, 2, sumfun)))
crit_sm$model = 'fixed thresholds'
crit0_sm$model = "random thresholds"
crit_sm$threshold = 1:7 - .1
crit0_sm$threshold = 1:7 + .1
df = rbind(crit_sm, crit0_sm)
names(df)[1:3] = c('lo', 'mean', 'hi')
ggplot(df, aes(threshold, mean, group = model, color = model)) + geom_point() + geom_line() + geom_errorbar(aes(ymin = lo, ymax = hi))
```

The point and interval estimates of the thresholds based on the
thresholds-as-"item-parameters" model are severely biased. In
particular, the uncertainty in the estimates is severely
underestimated in this model. Was it obvious to you that the model
with no threshold random effects is such a bad idea?

Now we will fit a model which is similar to an SDT model, but there
will be two important differences. We will not assume that the answer
"signal" cannot be less likely for the stimulus "signal" than for the
stimulus "noise", i.e., we will not assume that *d'* must be
non-negative, and we will *not* assume that the pattern of the
thresholds is the same for the "noise" and the "signal" stimuli. If
you know your Signal Detection Theory well, this may seem to you like
a perverse idea. If that is the case, please bear with me.

In this model each participant may have two different sets of
thresholds, one for the "noise" and one for the "signal" stimuli. We
will account for the fact that participant specific threshold effects
may be correlated; They are usually correlated, because if some
threshold is higher (lower) than the sample average, then the
threshold above (below) it may also be higher (lower).

Here we introduce the effect of the stimulus on the gamma / thresholds
vector. We use separate intercepts and slopes parametrization (-1 +
stim), i.e., the thresholds are to be estimated for each stimulus class
instead of estimating them for "noise" stimuli and estimating the
difference in the thresholds between "signal" and "noise".

```{r echo=T,results="show"}                
fixed1 = list(eta = ~ 1, gamma = ~ -1 + stim)
random1 = list(list(group = ~ id, gamma = ~ -1 + stim))
## We have to make sure that stim is a factor for the separate
## intercepts and slopes parametrization to work as intended.
d2$stim = as.factor(d2$stim)
## Note that because for an ordinal model the stimulus variable (here
## the 'stim' variable) is not special, we have to include it in the
## additional variables vector
adata1 = aggregate_responses(d2, response = 'r', variables = c('id', 'stim'))
sdata1 = make_stan_data(adata1, fixed1, random1, model = 'ordinal', gamma_link = link)
model1 = make_stan_model(random1, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit1 = stan(model_code = model1, data = sdata1,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit1, '~/windows/temp/ordinal_fit_1.rds')
}else{
    fit1 = readRDS('~/windows/temp/ordinal_fit_1.rds')
}
## We will not look at the eta parameter since it was fixed at 0
print(fit1, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
```

Now we can again translate between the gamma and the criteria vectors
(fixed effects) for each stimulus:

```{r echo=T,results="show"}
s1 = as.data.frame(fit1)
crit1 = gamma_to_crit(s1, 1, gamma_link = link)
crit2 = gamma_to_crit(s1, 1, beta_index = 2, gamma_link = link)
sumfun = function(x)c(quantile(x, c(.025, .975)), mean(x))[c(1,3,2)]
crit1_sm = apply(crit1, 2, sumfun)
crit2_sm = apply(crit2, 2, sumfun)
```

The criteria for the stimulus class "noise" look like this:

```{r echo=T,results="show"}
round(t(crit1_sm), 2)
```

and the criteria for the stimulus class "signal" look like this:

```{r echo=T,results="show"}
round(t(crit2_sm), 2)
```

Note the difference between the main criterion (row 4). This
difference approximates *d'*. Even though, as we will later see, this
model does not correctly account for the effect of the stimulus
(because it violates the non-negativity assumption), we can still
recover this effect pretty well, because *d'* is equivalent to a
constant shift of all the thresholds in one predefined
direction. Shifting all the criteria to the right is equivalent to
shifting the underlying distribution of latent values (or evidence) to
the left, so we have to subtract the average position of the criteria
for the "signal" stimuli from the average position of the criteria for
the "noise" stimuli. Here is our first attempt:

```{r echo=T,results="show"}
dprim1 = sumfun(s1[, 'gamma_fixed[4,1]'] - s1[, 'gamma_fixed[4,2]'])
round(dprim1, 2)
```

We will later see that this is our average *d'*, but it also isn't.

In an SDT model we usually assume that the stimulus class affects
*only* the *d'* parameter, and that both criteria (gamma) and *d'* may
differ between the participants (or items, etc.). In bhsdtr *d'* is
represented as *log(d') = delta*, which forces it to be
non-negative. This is an important assumption of Signal Detection
Theory - if this assumption is violated, for example, because of the
response reversal in some participants or conditions, then an SDT
model is not valid and a non-trivial generalization of SDT should be
used instead (see the bhsdtr preprint for a more detailed explanation
and some references).

```{r echo=T,results="show"}   
## We want one average d' / delta and one vector of average
## thresholds / criteria / gammma
fixed2 = list(delta = ~ 1, gamma = ~ 1)
## Each participant has a unique d' / delta and a unique set of
## thresholds / criteria / gamma
random2 = list(list(group = ~ id, delta = ~ 1, gamma = ~ 1))
```

Because of the importance of the stimulus class in SDT models
("noise", or 1, and "signal", or 2) we have to indicate which variable
represents this information ( aggregate_responses(..., stimulus =
'stim', ...) ) when calling the aggregate_responses function.

```{r echo=T,results="show"}
adata2 = aggregate_responses(d2, stimulus = 'stim', response = 'r', variables = c('id'))
sdata2 = make_stan_data(adata2, fixed2, random2, model = 'sdt', gamma_link = link)
model2 = make_stan_model(random2, model = 'sdt', gamma_link = link)
if(fresh_start){
    fit2 = stan(model_code = model2, data = sdata2,
                pars = c('delta_fixed', 'gamma_fixed', 'delta_sd_1', 'gamma_sd_1',
                         'delta_random_1', 'gamma_random_1',
                         ## We will need this for the plot_sdt_fit function
                         'counts_new'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit2, '~/windows/temp/ordinal_fit_2.rds')
}else{
    fit2 = readRDS('~/windows/temp/ordinal_fit_2.rds')
}
print(fit2, probs = c(.025, .975), pars = c('delta_fixed', 'gamma_fixed', 'delta_sd_1', 'gamma_sd_1'))
```

So how well do the two *d'* estimates match? We can translate between
the delta and the *d'* parameters by using exponentiation. We have to
do this first, before doing anything else with the posterior samples,
because the exponential function is non-linear.

```{r echo=T,results="show"}
s2 = as.data.frame(fit2)
dprim2 = sumfun(exp(s2[,'delta_fixed[1,1]']))
round(cbind(dprim1, dprim2), 2)
```

The estimates match poorly. Or do they? What about the individual *d'*
estimates? To calculate the individual shifts in the thresholds /
criteria in the more general ordinal model we have to add the random
effects of participants to the fixed effect:

```{r echo=T,results="show"}   
dprim1i = (s1[['gamma_fixed[4,1]']] + s1[, grep('gamma_random_1\\[.+,4,1]', names(s1))]) -
    (s1[['gamma_fixed[4,2]']] + s1[, grep('gamma_random_1\\[.+,4,2]', names(s1))])
dprim1i = as.data.frame(t(apply(dprim1i, 2, sumfun)))
dprim2i = as.data.frame(t(apply(exp(s2[['delta_fixed[1,1]']] +
                                    s2[, grep('delta_random_1', names(s2))]), 2, sumfun)))
names(dprim1i) = names(dprim2i) = c('lo', 'mean', 'hi')
dprim1i$model = 'ordinal'
dprim1i$i = 1:nrow(dprim1i) - .1
dprim2i$model = 'sdt'
dprim2i$i = 1:nrow(dprim2i) + .1
ggplot(rbind(dprim1i, dprim2i), aes(i, mean, group = model, color = model)) +
    geom_point() + geom_errorbar(aes(ymin = lo, ymax = hi)) +
    geom_abline(intercept = 0, slope = 0) +
    ylab('d\'') +
    labs(title = sprintf('Correlation between individual d\' estimates: r = %.2f',
                         cor(dprim1i$mean, dprim2i$mean)))
```

The point and interval estimates match quite well (*r* = .93 for the
point estimates), with some notable exceptions. According to the
ordinal model some credible intervals around individual *d'* point
estimates include 0 and one point estimate is even below 0, which
contradicts an important assumption of the SDT model. We will address
this issue later.

If the participant-specific *d'* estimates match, the average *d'*
should also match, right? Maybe the SDT model does not fit the data
well?

```{r echo=T,results="show"}   
plot_sdt_fit(fit2, adata2, c('id'), type = 'response', bw = F, verbose = F)
```

The SDT model seems to fit very well in a sense that the observed
distributions of responses are well within the 95% posterior
predictive intervals for each participant. However, this does not mean
much; An SDT model with multiple criteria is not saturated, so it does
not have to fit the data, but it does not leave too many degrees of
freedom in the data either. We can safely assume a priori that it will
usually fit the data well, regardless if it is true or not.

The discrepancy between the average *d'* estimates from both models is
not really a consequence of allowing for negative *d'* values in the
ordinal model. This discrepancy is a consequence of the assumed
distribution of individual shifts / *d'* values. In the ordinal model
we assumed that the shifts (*= d'*) are normally distributed, whereas
in the SDT model we assumed that *log(d')* are normally
distributed. We can see if the assumption that *log(d')* (or *d'*) are
normally distributed is approximately correct in our dataset. First,
let's take a look at the participant-specific delta estimates in the
SDT model:

```{r echo=T,results="show"}
## Participant's delta = average delta + participant's effect
delta2 = apply(s2[['delta_fixed[1,1]']] + s2[, grep('delta_random_1', names(s2))], 2, mean)
ggplot(data.frame(delta = delta2), aes(sample = delta)) +
    stat_qq() +
    ylab('Sample quantile of log(d\') = delta random effects') +
    xlab('Theoretical normal quantile') +
    labs(title = 'SDT model')
```

The individual *delta = log(d')* estimates in the SDT model seem to be
approximately normally distributed. We cannot plot *log(d')* estimates
based on the ordinal model, because according to this model one
participant has a negative *d'* value and even without this
participant's data there are quite a few negative posterior *d'*
samples. However, we can see if the *d'* point estimates based on the
general ordinal model are approximately normally distributed:

```{r echo=T,results="show"}   
ggplot(data.frame(dprim = dprim1i$mean), aes(sample = dprim)) +
    stat_qq() +
    ylab('Sample quantile of d\' random effects') +
    xlab('Theoretical normal quantile') +
    labs('Ordinal model')
```

They seem to be approximately normally distributed as well. The
average *d'* in each model is estimated as a mean of different
distributions: in the ordinal model participant-specific shifts are
assumed to be normally distributed, but in the SDT model *delta =
log(d')* random effects are assumed to be normally distributed. We can
easily account for this by making use of the relationship between the
parameters of the normal distribution and the associated log-normal
distribution:

```{r echo=T,results="show"}   
round(cbind(sumfun(exp(s2[['delta_sd_1[1,1]']]^2 / 2 + s2[['delta_fixed[1,1]']])),
            dprim1), 2)
```

Now we see that the two models give very similar point and interval
estimates of average *d'*. How cool is this?

To summarize, the SDT model is much simpler (it has only one set of
thresholds for each participant), it fits the data well at the
individual level, and it correctly assumes that the individual *d'* /
shift values are log-normally distributed. SDT is a clear winner,
right? Not yet it isn't.

# Causal assumptions of Signal Detection Theory

An SDT model is not only a statistical model, it also has *causal*
assumptions. To be clear, an SDT model *as a statistical model* does
not have any causal assumptions. However, causal assumptions are
present when the parameters of this model are interpreted as
properties of a psychological process, or when it is used to
*deconfound* (a causal notion) sensitivity (another extra-statistical
notion) from bias (this term has a technical meaning in statistics,
but it means something *psychological* in an SDT model - *a tendency
to prefer one class of stimuli over the other*).

Because the gabor dataset comes from a study in which the stimulus
class was chosen at random on every trial, every estimate of the
statistical effect of the stimulus class on any parameter of the model
is an unbiased estimate of the total causal effect of the stimulus
class on this parameter. Moreover, in such designs every statistical
effect of the participant factor is an estimate of the property or the
causal effect of the participants. This means that both the individual
*d'* parameters and the individual thresholds are participant-, not
item-, parameters in this model. So far so good.

However, in the SDT model we also assumed that the stimulus class can
*only* affect the distance between the evidence distribution means (=
*d'*), whereas in the ordinal model we allowed for the possibility of
the stimulus class having an effect on the pattern of the
thresholds. It may seem odd to you to allow for this, but there is
nothing in the design of the study that justifies the causal
assumption of selective influence. The estimates of the thresholds are
based on the data that were generated after the stimulus was
presented, and so, theoretically, the pattern of the thresholds could
be affected by the stimulus class as well.

It does not matter if an SDT model with selectivity assumption fits
the data well, because if the effect of the stimulus class is not
selective, the model is simply false. In particular, when the
selectivity assumption is false, an SDT model cannot deconfound
sensitivity / *d'* from bias / threshold / criteria effects. Once we
drop the selectivity assumption there is no longer any *d'* / internal
value parameter, only the thresholds, like in our general ordinal
model. The selectivity assumption is central to Signal Detection
Theory. If you know your Signal Detection Theory well, please take a
moment and try to recall one instant of this important assumption
being explicitly addressed in a paper or a book on Signal Detection
Theory. If you manage to find one such example, please send me an
email.

Before we go any further, just for fun, let's see what happens if we
introduce yet another assumption of selective influence. We will fit a
hierarchical SDT model in which we will treat the thresholds as
"item-parameters" and we will treat the *d'* parameter as a
"person-parameter".

```{r echo=T,results="show"}
fixed3 = list(delta = ~ 1, gamma = ~ 1)
## Each participant has a unique d', but not thresholds
random3 = list(list(group = ~ id, delta = ~ 1))
adata3 = aggregate_responses(d2, stimulus = 'stim', response = 'r', variables = c('id'))
sdata3 = make_stan_data(adata3, fixed3, random3, model = 'sdt', gamma_link = link)
model3 = make_stan_model(random3, model = 'sdt', gamma_link = link)
if(fresh_start){
    fit3 = stan(model_code = model3, data = sdata3,
                pars = c('delta_fixed', 'delta_sd_1', 'gamma_fixed', 'delta_random_1',
                         ## We will need this for the plot_sdt_fit function
                         'counts_new'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit3, '~/windows/temp/ordinal_fit_3.rds')
}else{
    fit3 = readRDS('~/windows/temp/ordinal_fit_3.rds')
}
print(fit3, probs = c(.025, .975), pars = c('delta_fixed', 'gamma_fixed', 'delta_sd_1'))
```

Does the model fit?

```{r echo=T,results="show"}
plot_sdt_fit(fit3, adata3, type = 'response', bw = F, verbose = F)
```

At the aggregated data level the fit is not perfect, but maybe it is
not that terrible either. What about the fit at the individual level?

```{r echo=T,results="show"}
plot_sdt_fit(fit3, adata3, 'id', type = 'response', bw = F, verbose = F)
```

A few points are outside the predictive intervals, but maybe that's
ok? What if we compared the point and interval estimates of the
so-called "person-parameters"?

```{r echo=T,results="show"}   
s3 = as.data.frame(fit3)
dprim3i = as.data.frame(t(apply(exp(s3[['delta_fixed[1,1]']] + s3[, grep('delta_random_1', names(s3))]), 2, sumfun)))
names(dprim3i) = c('lo', 'mean', 'hi')
dprim3i$model = 'person- and item-parameters'
dprim3i$i = 1:nrow(dprim3i) - .1
ggplot(rbind(dprim2i, dprim3i), aes(i, mean, group = model, color = model)) +
    geom_point() + geom_errorbar(aes(ymin = lo, ymax = hi)) +
    geom_abline(intercept = 0, slope = 0) +
    labs(title = sprintf('Correlation between individual d\' estimates: r = %.2f', cor(dprim2i$mean, dprim3i$mean)))    
```

This does not look good at all. In particular, some of the point and
interval estimates are clearly biased. If we compare the average width
of the intervals...

```{r echo=T, results="show"}
round(mean(dprim3i$hi - dprim3i$lo) / mean(dprim2i$hi - dprim2i$lo), 2)
```

... we see that the simplified model not only gives biased
"person-parameter" estimates but the estimates seem more reliable
(they are shorter by about 20% on average). It does not make sense to
look at the estimates of the "item-parameters" since we already know
that they have little in common with the true thresholds.

Now back to Signal Detection Theory and the assumption of selective
influence. We can at least try to test this assumption. We will not
alter the ordinal model code to account for the non-negativity of *d'*
(and the log-normal distribution of random *d'* effects) just yet. As
we already know, even without this correction the estimates of
participants' shifts are similar to the *d'* estimates.

Note that we do not use the separate intercepts and slopes
parametrization here (we use ~ stim instead of ~ -1 + stim), because
we are very much interested in the difference in the pattern of the
thresholds caused by the stimulus class.

```{r echo=T,results="show"}   
## Separate thresholds for "signa" and "noise"
fixed4 = list(eta = ~ 1, gamma = ~ stim)
## Each participants has two unique sets of thresholds, one for "signa" and one for "noise".
random4 = list(list(group = ~ id, gamma = ~ stim))
adata4 = aggregate_responses(d2, response = 'r', variables = c('id', 'stim'))
sdata4 = make_stan_data(adata4, fixed4, random4, model = 'ordinal', gamma_link = link)
model4 = make_stan_model(random4, model = 'ordinal', gamma_link = link)
if(fresh_start){
    fit4 = stan(model_code = model4, data = sdata4,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit4, '~/windows/temp/ordinal_fit_4.rds')
}else{
    fit4 = readRDS('~/windows/temp/ordinal_fit_4.rds')
}
print(fit4, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
```

Here we see the estimated average shift / *d'* thought of as the mean
of the normally distributed *d'* (not *delta = log(d')*) random
effects again:

```{r echo=T,results="show"}
s4 = as.data.frame(fit4)
round(cbind(sumfun(-s4[['gamma_fixed[4,2]']]), dprim1), 2)
```

Ok, we did simplify things a bit, but it still makes sense to take a
look at the other gamma effects, because these other effects
correspond to the variability in the *pattern* of the thresholds, and
the estimates of the thresholds in our ordinal model seem perfectly
fine.

```{r echo=T,results="show"}   
round(HPDinterval(as.mcmc(s4[, grep('gamma_fixed', names(s4))])), 2)
```

Among the parameters that represent the stimulus-induced average
differences between the gamma parameters (i.e., gamma_fixed[.,2]) the
gamma_fixed[4,2] parameter does not seem to be the only one that is
non-zero. This indicates that the pattern of the thresholds - not only
their average position (= gamma_fixed[4,.]) - may be affected by the
stimulus class. Interestingly, the random effects' standard deviations
tell a different story:

```{r echo=T,results="show"}   
round(HPDinterval(as.mcmc(s4[, grep('gamma_sd_1', names(s4))])), 2)
```

Note the clear pattern for the gamma_sd_1[.,2] parameters. The
*average* pattern of the thresholds seems to be different for the two
stimulus classes (gamma_fixed[.,2] except for gamma_fixed[4,2] ~ d'),
and participants' thresholds for the "noise" stimuli seem to differ
(gamma_sd_1[.,1]) - none of the 95% HPD intervals for the
gamma_sd_1[.,1] parameters include zero: in terms of standard
deviations of posterior distributions of standard deviations (phew)
most are in fact quite far from zero. However, the difference in the
pattern of the thresholds for the "signal" and the "noise" stimuli
seems to be fairly constant across the participants: all the
gamma_sd_1[.,2] parameters seem to be near-zero, except for the
gamma_sd_1[4,2] parameter, which corresponds to the variability in the
*d'* parameter. This is weird.

We will modify the stan model code to account for the non-negativity
of *d'*. Let's index the rows of our dataset by *i*. Because of the
way the model was parametrized...

```{r echo=T,results="show"}   
head(sdata4$X_gamma)
```

... when stim = "noise" = 0, gamma[4]_i = gamma_fixed[4,1] +
gamma_random_1[g_i][4,1], where g_i is the participant index in row
*i*. We can leave this part of the model intact, since this is the
main criterion position, which is an unconstrained parameter, but
also, as we can see below, the observed distribution of the main
criterion random effects is approximately normal:

```{r echo=T,results="show"}
ggplot(data.frame(gamma_4 = apply(s4[['gamma_fixed[4,1]']] + s4[, grep('gamma_random_1\\[.+,4,1]', names(s4))], 2, mean)),
       aes(sample = gamma_4)) +
    stat_qq() +
    ylab('Sample quantile of main criterion random effects') +
    xlab('Theoretical normal quantile') +
    labs('Ordinal model')
```

When stim = "signal" = 1, gamma[4]_i = gamma_fixed[4,1] +
gamma_random_1[g][4,1] + gamma_fixed[4,2] + gamma_random_1[g][4,2]. We
want this other part (the last sum which represents the difference in
gamma[4] between the "noise" and the "signal" stimuli) to represent
log of average distance between the thresholds = log(shift) = log(d'),
so we have to subtract it from gamma[4]_i, exponentiate this sum, and
add it to gamma[4]_i.

```{r echo=T,results="show"}   
fixed5 = list(eta = ~ 1, gamma = ~ stim)
random5 = list(list(group = ~ id, gamma = ~ stim))
adata5 = aggregate_responses(d2, response = 'r', variables = c('id', 'stim'))
sdata5 = make_stan_data(adata5, fixed5, random5, model = 'ordinal', gamma_link = link)
model5_ = make_stan_model(random5, model = 'ordinal', gamma_link = link)
## Here we introduce the necessary change in the model code. Kb2
## stands for K by 2 = the middle criterion index.
model5 = gsub('criteria\\[Kb2] = gamma\\[Kb2];',
              'if(X_gamma[n,2] == 0){
                  criteria[Kb2] = gamma[Kb2];
               }else{ 
                  criteria[Kb2] = gamma[Kb2] - (gamma_fixed[Kb2,2] + gamma_random_1[group_1[n]][Kb2,2]) - exp(gamma_fixed[Kb2,2] + gamma_random_1[group_1[n]][Kb2,2]);
               }',
              model5_)
```

This way the gamma[4,1] parameter represents the position of the main
criterion and gamma[4,2] = delta = log(d').

```{r echo=T,results="show"}   
if(fresh_start){
    fit5 = stan(model_code = model5, data = sdata5,
                pars = c('gamma_fixed', 'gamma_random_1', 'gamma_sd_1'),
                init_r = .5,
                iter = 5000,
                chains = 4)
    saveRDS(fit5, '~/windows/temp/ordinal_fit_5.rds')
}else{
    fit5 = readRDS('~/windows/temp/ordinal_fit_5.rds')
}
print(fit5, probs = c(.025, .975), pars = c('gamma_fixed', 'gamma_sd_1'))
```

Now let's have a look at the *d'* estimate (fixed effect) and compare
it with the *d'* estimate in the SDT model:

```{r echo=T,results="show"}
s5 = as.data.frame(fit5)
round(cbind(sumfun(exp(s5[['gamma_fixed[4,2]']])), dprim2), 2)
```

This is close. Not the same, but the model is different. What about
the bias = main criterion relative the the midpoint between the main
criteria?

```{r echo=T,results="show"}   
round(cbind(sumfun(s5[['gamma_fixed[4,1]']] - exp(s5[['gamma_fixed[4,2]']]) / 2),
      sumfun(s2[['gamma_fixed[4,1]']])), 2)
```

Also close.

We have fitted a hierarchical SDT model which respects the
non-negativity assumption, the ordering of the thresholds, accounts
for the log-normal distribution of the *d'* random effects, allows for
random effects in both the *d'* and the individual threshold
parameters while still respecting the SDT assumptions (thanks to an
order-preserving gamma link function and the delta log link function
which respects the non-negativity assumption), and accounts for the
correlations between the random effects. First such model was
described in the bhsdtr preprint. This model, however, does not assume
that the stimulus class affects only the *d'* parameter. We can now
again see if there is any evidence of non-selectivity:

```{r echo=T,results="show"}   
round(HPDinterval(as.mcmc(s5[, grep('gamma_fixed', names(s5))])), 2)
```

```{r echo=T,results="show"}                
round(HPDinterval(as.mcmc(s5[, grep('gamma_sd_1', names(s5))])), 2)
```

We see again that the gamma_sd_1[.,2] parameters, except for
gamma_sd_1[4,2] = log(d'), may be zero - the effect of the stimulus on
all the log-distances seems to be more or less the same for each
participant. What is going on here?

Interaction is tricky. For example, if you have an additive effect of
two factors on the original scale of the dependent variable...

```{r echo=T,results="show",fig.asp=.6}   
df = expand.grid(f1 = 1:2, f2 = 1:2)
df$y = df$f1 + df$f2
df$f1 = as.factor(df$f1)
df$f2 = as.factor(df$f2)
ggplot(df, aes(f1, y, group = f2, color = f2)) + geom_line()
```

...then after applying a non-linear transformation to the dependent
variable you may observe a significant interacive effect:

```{r echo=T,results="show",fig.asp=.6}
ggplot(df, aes(f1, exp(y), group = f2, color = f2)) + geom_line()
```

This means that whenever you see an interactive effect it may be an
artefact of the scale of the dependent variable. So, instead of
examining the effects in the gamma parameters. let's have a look at
the average *criteria / thresholds*. Notice that we have to add the
effect on gamma to gamma for stimulus = "noise" to obtain gamma for
stimulus = "signal".

```{r echo=T,results="show"}   
crit1 = gamma_to_crit(s5, gamma_link = link)
crit2 = gamma_to_crit(s5[, grep('gamma_fixed\\[.+,1]', names(s5))] + s5[, grep('gamma_fixed\\[.+,2]', names(s5))], gamma_link = link)
```

Now we can directly test if the average pattern of the criteria may be
different by comparing the *centered* criteria vectors:

```{r echo=T,results="show"}
round(HPDinterval(as.mcmc((crit2 - apply(crit2, 1, mean)) - (crit1 - apply(crit1, 1, mean)))), 2)
```

When we center the criteria for the two classes of stimuli we no
longer see any evidence of the differences in their relative
positions. The only thing that seems to differ, and not by much, is
the main criterion, but this, I think, is confounded with *d'* in our
model (this is only my guess). Selectivity seems to hold, kind of. To
really justify the selectivity assumption we have to argue for the
null, which, in a bayesian setting, usually means using the bayes
factor. Even though I love bayesian inference, I do not like the bayes
factor, and I am not alone in this.

This is one way to *begin* to test an SDT model. An SDT model has
other problematic assumptions. One is that the time-accuracy tradeoff
can be ignored. If participants or conditions differ in the response
time, then the differences in *d'* may not mean that sensitivity is
different, because participants may simply take more time to respond
in certain experimental conditions, which inflates the *d'*
estimate. You need something like the diffusion model to deal with
this problem. If you combine the diffusion model with ratings, you
also need something like a hierarchical diffusion model with ratings
and an order-preserving link function.

Among the other problems that need to be addressed some have to do
with non-linearity and aggregation. If you fit an SDT model, or any
other non-linear model, to data aggregated over grouping factors, such
as participants or items, because of non-linearity, which does not
play well with aggregation, some or all of the point and interval
estimates may be biased, perhaps severely so, and the inference will
be invalid. You can see how horribly biased the point *and* interval
estimates of every SDT parameter may become when you overly aggregate
your data in the bhsdtr preprint. By the way, I am now in the process
of examining the direction and magnitude of bias in SDT models fitted
to overly-aggregated data by using the excellent Confidence database
created by Doby Rahnev.

All this means that every SDT analysis to this day that was done using
data aggregated over participants is invalid. In particular, to my
knowledge, to this day there is little *valid* evidence that the
Unequal Variance SDT model is closer to the true model than the Equal
Variance SDT model.

Note also that here we were aggregating over trials, which means that,
if the SDT parameters may differ between trials, and they certainly do
differ, the only question is by how much, then all the SDT estimates
that I have presented to you are biased to an unknown extent. There is
a lot to be done before we can say how well does the humble SDT model
deconfound sensitivity from bias in any particular case.

What if we tried this trick with testing the selectivity assumption,
but using the Unequal Variance SDT model, or the meta-d' model, or
some other generalization of the SDT model? We cannot introduce any
new parameters to our model, because it is already saturated. What we
can do, for example, is fit an Unequal Variance SDT model, estimate
the ratio of the variances, fix the ratio, introduce it to the
saturated ordinal model, thus turning it into the Fixed Unequal
Variance Ordinal Model. I think this is a neat idea. If there is
evidence of non-selectivity in this model, we can keep trying with
other modifications. If non-selectivity disappears, we have evidence
that the selectivity assumption may be correct *and* that the Unequal
Variance SDT model may be better than the Equal Variance SDT model.

# A rant about Item Response Theory (from someone who is relatively new to Item Response Theory)

An SDT model is essentially an ordinal model with some additional
assumptions (e.g., selectivity, non-negativity of d', no time-accuracy
tradeoff issues, etc.). When there is more than one criterion /
threshold this model represents the distribution of polytomous ordinal
responses. Because by definition a polytomous item has a multinomial
distribution and the multinomial distribution does not have a separate
parameter for the residual variance, when item-parameter variability
is not accounted for by the model it can only be hidden or misplaced
(overdispersion). The correct way to account for item-parameter
variability in ordinal polytomous items is to allow for individual
threshold effects while respecting the ordering of the thresholds. In
the ordinal model that we consider here this can only be achieved by
using the order-preserving link functions.

Assume that the ordered responses ranging from 1 to K come from the
same general kind of process as the responses in an SDT model when the
stimulus is 1 ("noise"). For example, when John is faced with the
Likert-type item "How often did you cry during the last two weeks" an
unknown process produces some "internal experience" value. This latent
value manifests itself in the response "never" (= 1) if it is low
enough, "a few times" (= 2) if it is a bit higher than some latent
threshold, or in the response "frequently" (= 3) if it is higher than
some other, still higher, threshold. The thresholds are assumed to be
ordered, because otherwise, the item would not have the intended
meaning and thus would not be a valid instrument of measuring
depression.

Now consider a participant who repeatedly responds to the same item
and who has the memory of this event wiped out after each
response. Despite the wiping out of memory we can still expect some
variability in the responses due to the instability of the complex
process which produces the internal value, as well as due to the
instability of the decision criteria. This variability is certainly
there, the question is if it is non-negligible, and if so, what can be
done about it.

Unfortunately, without additional assumptions, it is impossible to
model both kinds of variability at the same time even in the
hypothetical repeated-measurement-with-memory-wipeout-scenario,
because every shift in the internal value is indistinguishable from a
uniform effect in the thresholds, i.e., a constant shift of all the
thresholds. As I hope you already realize, the required additional
assumptions that make such models identifiable under certain
conditions are necessarily causal.

In many situations, this is both a statistical and a conceptual
problem. In particular, when the design is purely observational -
e.g., questionnaire measurement - without additional causal
assumptions, the way of describing the process of responding to a
Likert-type item in terms of internal values and thresholds is simply
redundant. As far as the design of the study is concerned, all we can
say about the internal value is that it is relative to the thresholds-
and the position of each threshold is relative to the positions of the
remaining thresholds and to the position of the internal
value. Without the selectivity assumption, to say that John thinks
that he was not crying at all during the last two weeks and to say
that *John's* threshold for the response "a few times" is above the
amount of crying that John thinks he experienced, or decided to
reveal, is to say the same thing.

It follows that *in such situations, we do not need the internal value
parameter at all*, nor should we introduce it without a very good
reason to do so. That is, unless there is some factor about which we
can perhaps assume that it can *only* produce a uniform effect in the
thresholds, such as the stimulus class ("noise" or "signal") in a
binary classification task, which can *perhaps* be assumed to induce a
change (*d'*) in the internal value only; In such cases the internal
value parameter may be justified and useful. We have obtained evidence
that this assumption perhaps holds for the analyzed dataset.

So what about the central promise of Item Response Theory? Isn't it
true that we can separate the internal values ("person parameters")
from the thresholds ("item parameters") and place them on a common
scale? Once the participants leave the room the items retain their
estimated "difficulty levels", right? The item difficulty (=
thresholds) is not defined relative to some sample of participants who
filled the questionnaire. How is this separation possible if the two
kinds of effects are completely confounded? In particular, how does
Item Response Theory achieve this magical separation when *each person
provides only a few responses, usually just one, to each item*?

It is usually not an *achievement* of Item Response Theory, it is just
an unrealistic *causal assumption*. The illusory nature of this
achievement becomes obvious as soon as we realize that without
additional strong causal assumptions it is impossible to distinguish
between genuine internal value effects and (dis)simulation (i.e.,
faking), which conceptually seem to correspond to the change in the
*participant*-specific, not item-specific, decision thresholds, also
known as "item-parameters".

There is nothing in the design of the process of measurement by a
questionnaire that makes it possible to distinguish between the
genuine differences in internal values and the differences in the way
of responding. Because of random assignment of the stimulus class
above chance performance in a binary classification task cannot be
faked, and so the *d'* parameter definitely captures at least part of
the effect of the stimulus on the internal value. On the other hand,
every possible pattern of responses to a questionnaire can be faked.

In the context of Item Response Theory the issue of measurement
non-invariance due to the variability in the "item-parameters" is
often reduced to the problem of Differential Item Functioning
(DIF). However, the variability in "item-parameters" that I talk about
here has nothing to do with DIF, because DIF is a *population* level
effect; An estimate of DIF is obtained when the estimates of
"item-parameters" obtained for two different (often large) samples of
participants are compared. Moreover, these estimates are obtained by
more or less arbitrarily fixing the "item-parameters" within each
sample.

If we campared the estimates of the main decision criterion (i.e.,
bias) in SDT models fitted to two different large *samples* of
participants the difference would likely be negligible with very
narrow confidence / credible intervals, even though *individual
participants* clearly differ in how they use the main decision
criterion. By the same token, variability in item parameters in IRT
models cannot be correctly estimated just by assuming selective
influence and comparing estimates of "item-parameters" obtained in two
different samples, because estimates of sample average item-parameters
*hide or misplace* (overdispersion) the intra- or inter-individual
variability in item parameters.

By assuming that internal values are participant-specific and item
parameters are item-specific we force the estimates of item parameters
to become estimates of average person-specific thresholds. This not
only begs the question, but also, because the model is non-linear,
makes the estimates of item-parameters *asymptotically biased*. It
does not take much thought to see that the assumption of constant
"item-parameters" at the group / sample / population level is utterly
unrealistic. It follows that the achievement of separating participant
and item effects in IRT models is often not an achievement at all, it
is just wishful thinking.

The primary justification of the selective influence assumption comes
from a psychological theory of the task, not from the design of the
study. All that is guaranteed by randomly assigning the stimulus class
(i.e., by choosing the stimulus class at random on every trial) is
that any statistical effect of the stimulus class is an unbiased
estimate of the total causal effect of the stimulus class. To
*empirically* justify the assumption of *selective* influence of the
stimulus class on the internal evidence value we have to *test* if the
stimulus class does not also affect the thresholds. In principle, this
means that we have to argue for the null.

There are a few important advantages that an SDT model fitted to
binary classification task data has over an IRT model fitted to
questionnaire data. One is that the causal assumption of selective
influence of the stimulus class on the internal values is much more
plausible than the - implicit in many applications of IRT - causal
assumption of selective influence of participants on the
"person-parameters". The other is that in a typical binary
classification task case the process of responding is repeated many
times, the task is relatively simple ("was the gabor patch tilted left
or right" vs "do you wish you were as happy as other people seem to
be?") and the processes responsible for the generation of the internal
values and for the placement of the decision thresholds are likely to
be more stable.

The thresholds are determined by the items, by the properties of the
participants, and by the context in which the measurement takes
place. The assumption that an important latent aspect of the process
of psychological measurement is constant across participants is so
ridiculous it should be laughed at on the street.

# The missing part

At the risk of sounding ridiculous, I will now make use of the fact
that I can say whatever I want in an informal tutorial such as this
one. Also, I just had a beer.

It may seem that once (maybe in some distant future) we solve the
statistical problems and the causal problems of testing an SDT model
we can say that we are done. We have the great formal tools of
statistical and causal inference at our disposal, and they are
indispensable, but there is one last layer of abstraction that we did
not explicitly identify.

Let's assume that both the causal and the statistical assumptions of
an SDT model are true. This means that *d'* represents the total
causal effect of the stimulus class on the distribution of the
responses. However, both kinds of assumptions may be true about an
arbitrary physical process. For example. these assumptions may be true
about some diagnostic device. We can say that this device is more or
less sensitive, but the meaning of the term "sensitivity" is different
in an important way than when we say that the participant *who
performs the task* of binary classification is sensitive to the class
of the stimulus.

The meaning is different, because an arbitrary device is not yet an
autonomous purposeful agent. In particular, the sense in which an
arbitrary device "does" something i different than the sense in which
a participant does something. To be a bit more precise, an arbitrary
device does not *try* to do something, it just so happens that it was
designed by an autonomous purposeful agent (one of us) to solve the
problem faced by that agent, but it is the agent who "does all the
trying". This layer of abstraction has to do with behavior, purpose,
and function.

It is only at this layer of abstraction that a mechanism becomes a
*psychological* mechanism. It is my belief that it may be possible to
formalize the process of inference at this level of abstraction. If
such a formal theory of psychological inference was to be discovered,
it would perhaps resemble an abstract theory of agents, tasks, and
cognitive functions. Rational analysis (J.R. Anderson) and
probabilistic models of cognition (Chater, Oaksford, Tenenbaum,
Griffiths, etc) seem to be promising starting points, as does the
Reinforcement Learning framework (Sutton and Barto) in which, to my
knowledge, the first promising weak (a good thing), formal definition
of purposeful action (i.e., of behavior) was proposed, inspired by the
writings of Tolman and Perry.

It is no accident that the internal values are often called "evidence
samples" in Signal Detection Theory, or that the main criterion is
interpreted in terms of bias, or that there are deep smilarities
between the diffusion model and the process of sequential analysis. I
hope I have made the point that explicitly addressing not only the
statistical, but also the causal assumptions of Signal Detection
Theory may dramatically improve our understanding of the process of
fitting and interpreting the parameters of an SDT model. Who knows
what we could achieve if the statistical, causal, *and* functional
layers were explicitly combined.